---
title: "Planning"
format: html
---

Goal: 

Document for outlining the loop code to get a design-based estimate of variance from point transect distance sampling simulation.

Notes: 
  - our survey means the survey and analysis from the theoretical density, theoretical population, and the survey we conduct on that population

# Packages

```{r, message= F}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
```

Step 1

- based mostly on https://distancesampling.org/dsims/articles/GettingStarted.html 

- simulate a density surface over a square study area

  - should be 'lumpy' using 2D normal adjustments

- using dsims

Step 2

- produce a set of individual animal locations probabilistically from that density surface

- using dsims?

Step 3 

- estimate a density surface from that set of points

- using?

Step 4

- produce a set of detections from that density surface, analyze with half-normal detection ftn with no covariates

Step 5 

- Repeat step 4 for N simulations

# Theoretical Population

## Create survey object

This will be common to all analyses in this doc, just the survey grid

```{r}
# Create the survey region object
simsquare <- st_bbox(c(xmin = 0, xmax = 10000, ymin = 0, ymax = 10000)) |> 
  st_as_sfc() |> 
  st_as_sf()


region <- make.region(region.name = "simsquare",
                      shape = simsquare,
                      units = "m")
plot(region)
```

## Create theoretical density

Generating a density in this analysis is similar to the paper's method, that needs a density too.

This serves as the theoretical true density of animals in our study area

```{r}
# We first create a flat density grid
density_true <- make.density(region = region,
                        x.space = 500,
                        constant = 1)

# Now we can add some high and low points to give some spatial variability
density_true <- add.hotspot(object = density_true,
                       centre = c(2000, 8000),
                       sigma = 800,
                       amplitude = 1.2)

density_true <- add.hotspot(object = density_true,
                       centre = c(5000, 2000),
                       sigma = 6000,
                       amplitude = 0.5)

density_true <- add.hotspot(object = density_true,
                       centre = c(8000, 9000),
                       sigma = 1000,
                       amplitude = 2)

# Density plot
plot(density_true, region, scale = 0.001)
```

### Density GAM

optional step that smooths out the density using a GAM, not needed

boxlet smoothing instead of striplet smoothing

```{r}
# density_true.data <- density_true@density.surface[[1]]
# 
# fit.gam <- gam(density ~ s(x,y), data = density_true.data, family = gaussian(link="log"))
# 
# # Use the gam object to create a density object
# gam.density_true <- make.density(region = region,
#                             x.space = 500,
#                             fitted.model = fit.gam)
# 
# plot(gam.density_true, region, scale = 0.001)
```

## Population generation

generate the population from the density we just created

```{r}
pop.desc_true <- make.population.description(region = region,
                                        density = density_true,
                                        # density = gam.density_true,
                                        N = 5000,
                                        fixed.N = TRUE)

detect_true <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

pop_true <- generate.population(object = pop.desc_true, detect_true, region = region)

plot(pop_true, region)
```

Note: The next chunks (Design transects through run survey) have some redundancies to fitting the m1 models, this is because dsims machinery is expecting a simulation, we need those objects later. m1 generates an N_hat 

# Survey Design

## Design transects

points instead of lines, random starting point with a fixed angle.

```{r}
point.design_our_survey <- make.design(region = region, 
                            transect.type = "point",
                            design = "systematic",
                            spacing = 750,
                            edge.protocol = "minus",
                            design.angle = 65,
                            truncation = 750)

p.survey_our_survey <- generate.transects(point.design_our_survey)
plot(region, p.survey_our_survey)
```

## Specify detectability

we were unsure why we needed to give detectability

```{r}
# Create the detectability description
detect_our_survey <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

# Plot the simulation detection functions
plot(detect_our_survey, pop.desc_true)
```
## Specify ds() model

These set up the distance sampling analysis model for the theoretical survey

```{r}
analyses_our_survey <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             er.var = "P3",
                             criteria = "AIC")
```

## Set up simulation object

Sets up the simulation run that generates our survey sample

```{r}
sim.point_our_survey <- make.simulation(reps = 1,
                                design = point.design_our_survey,
                                population.description = pop.desc_true,
                                detectability = detect_our_survey,
                                ds.analysis = analyses_our_survey)
```

```{r}
# our_survey <- run.simulation(sim.point_our_survey)
```

## Run our survey

Runs the survey to get data points for the animals, representing the "true" sample

```{r}
our_survey <- run.survey(sim.point_our_survey)

plot(our_survey, region)
```


## Fit to theoretical survey data (ds())

fit to create the sim surface to estimate an abundance, and a sigma

```{r}
m1 <- ds(data = our_survey@dist.data,
         transect = "point",
         key = "hn",
         adjustment = NULL)

summary(m1)
```

stores estimates

```{r}
n_hat_our_survey <- m1$dht$individuals$N$Estimate

sigma_hat_our_survey <- coef(m1$ddf)
scaled_sigma_hat_our_survey <- exp(sigma_hat_our_survey$scale["(Intercept)", "estimate"])
```

# Create estimated density

Density surface estimation

uses some data wrangling to turn objects from our_survey, and from m1, into data objects that can be put into a dsm object.

fits a density surface model with dsm from those data by converting the model outputs from m1 into the surface.

```{r}
# dsm(count ~ s(x, y), ddf.obj = m1$ddf)
```

## Data wrangling

```{r}
our_survey_dist_data <- our_survey@dist.data

obsdata <- our_survey@dist.data |> # data from survey of our_survey
  filter(!is.na(distance)) |> # filters out non-detections 
  mutate(
    object = as.character(object), # converts detection label to character
    Sample.Label = as.character(Sample.Label), # converts individual label to character
    size = 1 # assigns group size 1, assumes individual animals
  ) |>
  select(object, Sample.Label, size, distance)
 
# all point transect locations

samplers <- our_survey@transect@samplers # pulls location of point transects
sampler_xy <- sf::st_coordinates(samplers) # puts x and y coords of each point into a two-column matrix
segdata <- samplers |>
  sf::st_drop_geometry() |>
  mutate(
    Sample.Label = as.character(transect),
    Effort = 1,
    x = sampler_xy[, "X"],
    y = sampler_xy[, "Y"]
  ) |>
  select(Sample.Label, Effort, x, y) # creates a dataset to be fed into dsm
                                      # includes sample label, effort, x, and y
```

## Fit density model

```{r}
# density surface model

dsm1 <- dsm(
  count ~ s(x, y),
  ddf.obj = m1,
  segment.data = segdata,
  observation.data = obsdata,
  family = quasipoisson(link = "log"),
  method = "REML",
  convert.units = 1) # uses ds model, segment data, and obsdata to fit a density surface to the observations from our survey
 
summary(dsm1)

class(dsm1)

plot(dsm1)
```
need abundance, density, detection ftn parameters

## Generate estimated density

This chunk uses the dsm1 model to create a new predicted density
  - this seems like a source of error that might contribute to the discrepancy between our calculated SEs from the boostrap and the fewster expectations
  

```{r}
pred_grid <- density_true@density.surface[[1]] |>
  mutate(area = as.numeric(sf::st_area(geometry))) 
# pulls the study area geometry from the true density as a scaffolding

pred_data <- sf::st_drop_geometry(pred_grid)
# adds the area of each cell to that scaffolding
 
pred_grid$Nhat <- predict(
  dsm1,
  newdata = pred_data,
  off.set = pred_grid$area,
  type = "response"
)
# predicts the abundance in each cell from the dsm1 model, which is based on our_survey
 
dsm_surface <- pred_grid |>
  mutate(density = pmax(Nhat / area, .Machine$double.eps)) |>
  select(strata, density, x, y, geometry)
# converts abundance into density
 
dsm_density <- make.density(
  region = region,
  x.space = 500,
  y.space = 500,
  density.surface = dsm_surface
)
# rebuilds a density object based on our model predictions
 
plot(dsm_density, region, scale = 0.001)
```

### status

now we have an estimated density surface from 'our survey' that comes from the theoretical true density and population

This density can now be used to create iterations of theoretical populations, these simulate the theoretical populations that can be generated during real analysis.

# Simulation framework

region is already created

density should come from dsm1:

```{r}
# # grid for dsims density object
# 
# grid_for_density <- density_true@density.surface[[1]] |>
#   select(x, y) |>
#   mutate(area = 500 * 500)
# # DSM predicts abundance per grid cell
# 
# grid_for_density$Nhat <- predict(
#   dsm1,
#   newdata = grid_for_density,
#   off.set = grid_for_density$area
# )
#  
# # convert cell abundance to density
# 
# grid_for_density <- grid_for_density |>
#   mutate(Dhat = Nhat / area)
#  
# # fit a plain mgcv GAM that predicts density from x,y only
# 
# dsm_density_gam <- mgcv::gam(
#   Dhat ~ s(x, y),
#   data = grid_for_density,
#   family = gaussian(link = "log"),
#   method = "REML"
# )
#  
# # now this can be passed to dsims
# 
# dsm_density <- make.density(
#   region = region,
#   x.space = 500,
#   fitted.model = dsm_density_gam
# )
# 
# # plot(dsm_density, region, scale = 0.001)
```

## Create dsims density

uses the density we generated with the dsm model to create a density object we can use in a dsims run
 
```{r}
# Use the dsm object to create a density object for dsims
sample.density <- make.density(region = region,
                            x.space = 500,
                            density.surface = dsm_surface)

plot(sample.density, region, scale = 0.001)
```

## Create dsims objects

Uses the sample objects 

```{r}
# Define the population description for the sample density
pop.desc <- make.population.description(region = region,
                                        density = sample.density,
                                        N = 5000,
                                        fixed.N = TRUE)
```

```{r}
# Define the point transect design for the sample density

point.design <- make.design(region = region,
                            transect.type = "point",
                            design = "systematic",
                            spacing = 750,
                            edge.protocol = "minus",
                            design.angle = 65,
                            truncation = 750)

p.survey <- generate.transects(point.design)
plot(region, p.survey)
```

```{r}
# Create the detectability description for the sample analysis
detect <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

# Plot the simulation detection functions
plot(detect, pop.desc)
```

```{r}
# create the analyses ds() specifications for the simulation runs

analyses <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             er.var = "P3",
                             criteria = "AIC")
```

## Set up e.g. simulation

```{r}
# Creates the simulation set-up

sim.point <- make.simulation(reps = 3,
                                design = point.design,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
```

```{r}
# Generate a single instance of a survey: a population, set of transects
# and the resulting distance data
eg.point.survey <- run.survey(sim.point)

# Plot it to view a summary
plot(eg.point.survey, region)
```

```{r}
sim.point <- run.simulation(sim.point)
```

```{r}
histogram.N.ests(sim.point)
```

```{r}
ests_vals <- sim.point@results$individuals$N

ests <- as.data.frame.table(ests_vals) |>
  pivot_wider(
    id_cols = c(Var1, Var3), # Stratum and Run
    names_from = Var2,       # Metric names (Estimate, se, cv, etc.)
    values_from = Freq       # The data values
  ) |>
  slice_head(n = 99) |>
  rename(region = Var1,
         sim = Var3,
         )
```

```{r}
sd(ests$Estimate)
```











