---
title: "Striplets_on_original_density"
format: html
---

# Packages

```{r}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
```

# Import Functions

```{r}
source(here::here("Functions.R"))
```

# Constants

```{r}
# true_N <- 5000
# 
# points_or_lines = "line"
 
trunc_dist <- 750
scale_parameter <- 200
design_spacing <- 750
 
density_grid_spacing <- 500
 
# design_angle <- 0
# bootstrap_reps <- 20
```

# Code

## Region

```{r}
region <- make_region(0, 10000, 0, 10000, "m")
```

## Density

```{r}
my_hotspots <- list(
  list(centre = c(2000, 8000), sigma = 800, amplitude = 1.2),
  list(centre = c(5000, 2000), sigma = 6000, amplitude = 0.5),
  list(centre = c(8000, 9000), sigma = 1000, amplitude = 2)
)

density_true <- make_hotspot_density(region, my_hotspots)
```

## prediction from density

```{r}
# Extract the true density surface and calculate expected N per cell
true_surface <- density_true@density.surface[[1]] |>
  mutate(
    area = as.numeric(sf::st_area(geometry)),
    N_hat_pred = density * area
  )

# Collapse to 1D (x-axis) for line transects
true_surface_1D <- true_surface |>
  group_by(strata, x) |>
  summarize(
    N_hat_pred = sum(N_hat_pred, na.rm = TRUE),
    area = sum(area, na.rm = TRUE),
    geometry = sf::st_union(geometry),
    .groups = "drop"
  ) |>
  mutate(density = pmax(N_hat_pred / area, .Machine$double.eps), y = 0) |>
  select(strata, density, x, y, N_hat_pred, area, geometry)
```

## Striplet code

```{r}
spacing <- design_spacing
sigma_hat <- scale_parameter
truncation <- trunc_dist

muvec <- true_surface_1D$N_hat_pred 
midvec <- true_surface_1D$x
musum <- sum(muvec, na.rm = TRUE)

bbox <- sf::st_bbox(region@region)
y_length <- as.numeric(bbox["ymax"] - bbox["ymin"])
x_min <- as.numeric(bbox["xmin"])
x_max <- as.numeric(bbox["xmax"])

B <- 50 
bvec <- seq(0, spacing, length.out = B)
Lbvec <- rep(0, B)
Abvec <- rep(0, B)

for (b_idx in 1:B) {
  b_val <- bvec[b_idx]
  lines_grid <- seq(x_min + b_val, x_max, by = spacing)
  Lbvec[b_idx] <- length(lines_grid) * y_length
  
  min_dist <- sapply(midvec, function(m) min(abs(m - lines_grid)))
  
  g_b <- exp(- (min_dist^2) / (2 * sigma_hat^2))
  g_b[min_dist > truncation] <- 0 
  
  Abvec[b_idx] <- sum(muvec * g_b, na.rm = TRUE)
}
    
var_ER_striplet <- mean((Abvec + (Abvec^2) * (1 - 1/musum)) / (Lbvec^2)) - (mean(Abvec / Lbvec))^2
```

since we dont have detections in this set-up as we are coming straight from the denisty, there is no other piece to the variance, it is just the encounter rate variance. Need to reason through what this means in isolation.

var_ER_striplet here is 209.6853

single fixed estimate, no variation since the density does not change. 

When run with the same settings, the var_ER_striplet in the main code, which factors in changing density estimates from the dsm() model is around: HAPPY TO EXPLORE THIS, WANTED TO CHECK BEFORE I DID SO















