---
title: "Planning"
format: html
---

# Goal

Document for outlining the loop code to get a design-based estimate of variance from point transect distance sampling simulation.

```{r, message= F}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(sf)
library(mgcv)
```

## Step 1

- based mostly on https://distancesampling.org/dsims/articles/GettingStarted.html 

- simulate a density surface over a square study area

  - should be 'lumpy' using 2D normal adjustments

- using dsims

## Step 2

- produce a set of individual animal locations probabilistically from that density surface

- using dsims?

## Step 3 

- estimate a density surface from that set of points

- using?

## Step 4

- produce a set of detections from that density surface, analyze with half-normal detection ftn with no covariates

## Step 5 

- Repeat step 4 for N simulations

## First Attempt

```{r}
# Create the survey region object

simsquare <- st_bbox(c(xmin = 0, xmax = 10000, ymin = 0, ymax = 10000)) |> 
  st_as_sfc() |> 
  st_as_sf()


region <- make.region(region.name = "simsquare",
                      shape = simsquare,
                      units = "m")
plot(region)
```

```{r}
# We first create a flat density grid
density <- make.density(region = region,
                        x.space = 500,
                        constant = 1)

# Now we can add some high and low points to give some spatial variability
density <- add.hotspot(object = density,
                       centre = c(2000, 8000),
                       sigma = 800,
                       amplitude = 1.2)

density <- add.hotspot(object = density,
                       centre = c(5000, 2000),
                       sigma = 6000,
                       amplitude = 0.5)

density <- add.hotspot(object = density,
                       centre = c(8000, 9000),
                       sigma = 1000,
                       amplitude = 2)

# Density plot
plot(density, region, scale = 0.001)
```

```{r}
density.data <- density@density.surface[[1]]

fit.gam <- gam(density ~ s(x,y), data = density.data, family = gaussian(link="log"))

# Use the gam object to create a density object
gam.density <- make.density(region = region,
                            x.space = 500,
                            fitted.model = fit.gam)

plot(gam.density, region, scale = 0.001)
```

```{r}
# Create a covariate list describing the distribution of cluster sizes
covariates <- list()

# Define the population description
pop.desc <- make.population.description(region = region,
                                        density = gam.density,
                                        covariates = covariates,
                                        N = 1000,
                                        fixed.N = TRUE)
```

```{r}
point.design <- make.design(region = region, 
                            transect.type = "point",
                            design = "systematic",
                            spacing = 1500,
                            edge.protocol = "minus",
                            design.angle = 65,
                            truncation = 750)

p.survey <- generate.transects(point.design)
plot(region, p.survey)
```

```{r}
# Create the detectability description
detect <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

# Plot the simulation detection functions
plot(detect, pop.desc)
```

```{r}
analyses <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             er.var = "P3",
                             criteria = "AIC")
```

```{r}
sim.point <- make.simulation(reps = 99,
                                design = point.design,
                                population.description = pop.desc,
                                detectability = detect,
                                ds.analysis = analyses)
```

```{r}
# Generate a single instance of a survey: a population, set of transects 
# and the resulting distance data
eg.point.survey <- run.survey(sim.point)

# Plot it to view a summary
plot(eg.point.survey, region)
```

```{r}
sim.point <- run.simulation(sim.point)
```

```{r}
histogram.N.ests(sim.point)
```

```{r}
ests_vals <- sim.point@results$individuals$N
              
ests <- as.data.frame.table(ests_vals) |>
  pivot_wider(
    id_cols = c(Var1, Var3), # Stratum and Run
    names_from = Var2,       # Metric names (Estimate, se, cv, etc.)
    values_from = Freq       # The data values
  ) |>
  slice_head(n = 99) |>
  rename(region = Var1,
         sim = Var3,
         )
```

```{r}
var(ests$Estimate)
```




















