---
title: "Planning"
format: html
---

# Goal

Document for outlining the loop code to get a design-based estimate of variance from point transect distance sampling simulation.

```{r, message= F}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
```

## Step 1

- based mostly on https://distancesampling.org/dsims/articles/GettingStarted.html 

- simulate a density surface over a square study area

  - should be 'lumpy' using 2D normal adjustments

- using dsims

## Step 2

- produce a set of individual animal locations probabilistically from that density surface

- using dsims?

## Step 3 

- estimate a density surface from that set of points

- using?

## Step 4

- produce a set of detections from that density surface, analyze with half-normal detection ftn with no covariates

## Step 5 

- Repeat step 4 for N simulations

## First Attempt

```{r}
# Create the survey region object

simsquare <- st_bbox(c(xmin = 0, xmax = 10000, ymin = 0, ymax = 10000)) |> 
  st_as_sfc() |> 
  st_as_sf()


region <- make.region(region.name = "simsquare",
                      shape = simsquare,
                      units = "m")
plot(region)
```

```{r}
# We first create a flat density grid
density_true <- make.density(region = region,
                        x.space = 500,
                        constant = 1)

# Now we can add some high and low points to give some spatial variability
density_true <- add.hotspot(object = density_true,
                       centre = c(2000, 8000),
                       sigma = 800,
                       amplitude = 1.2)

density_true <- add.hotspot(object = density_true,
                       centre = c(5000, 2000),
                       sigma = 6000,
                       amplitude = 0.5)

density_true <- add.hotspot(object = density_true,
                       centre = c(8000, 9000),
                       sigma = 1000,
                       amplitude = 2)

# Density plot
plot(density_true, region, scale = 0.001)
```

```{r}
density_true.data <- density_true@density.surface[[1]]

fit.gam <- gam(density ~ s(x,y), data = density_true.data, family = gaussian(link="log"))

# Use the gam object to create a density object
gam.density_true <- make.density(region = region,
                            x.space = 500,
                            fitted.model = fit.gam)

plot(gam.density_true, region, scale = 0.001)
```

### Animal position (poisson process)

```{r}
# Define the population description
pop.desc_true <- make.population.description(region = region,
                                        density = gam.density_true,
                                        N = 1000,
                                        fixed.N = TRUE)

detect_true <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

pop_true <- generate.population(object = pop.desc_true, detect_true, region = region)
```

### Generate the study grid and design

```{r}
point.design_our_survey <- make.design(region = region, 
                            transect.type = "point",
                            design = "systematic",
                            spacing = 1500,
                            edge.protocol = "minus",
                            design.angle = 65,
                            truncation = 750)

p.survey_our_survey <- generate.transects(point.design_our_survey)
plot(region, p.survey_our_survey)
```

### collect data

```{r}
# Create the detectability description
detect_our_survey <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

# Plot the simulation detection functions
plot(detect_our_survey, pop.desc_true)
```



```{r}
analyses_our_survey <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             er.var = "P3",
                             criteria = "AIC")
```

```{r}
sim.point_our_survey <- make.simulation(reps = 1,
                                design = point.design_our_survey,
                                population.description = pop.desc_true,
                                detectability = detect_our_survey,
                                ds.analysis = analyses_our_survey)
```

```{r}
# our_survey <- run.simulation(sim.point_our_survey)
```

```{r}
our_survey <- run.survey(sim.point_our_survey)
```

### Fit to those data

```{r}
m1 <- ds(data = our_survey@dist.data,
         transect = "point",
         key = "hn",
         adjustment = NULL) 

summary(m1)
```

```{r}
n_hat <- m1$dht$individuals$N$Estimate

sigma_hat <- coef(m1$ddf)
```

#### density surface estimation

```{r}
# dsm(count ~ s(x, y), ddf.obj = m1$ddf)
```

```{r}
obsdata <- our_survey@dist.data |>
  filter(!is.na(distance)) |>
  mutate(
    object = as.character(object),
    Sample.Label = as.character(Sample.Label),
    size = 1
  ) |>
  select(object, Sample.Label, size, distance)
 
# all point transect locations

samplers <- our_survey@transect@samplers
sampler_xy <- sf::st_coordinates(samplers)
segdata <- samplers |>
  sf::st_drop_geometry() |>
  mutate(
    Sample.Label = as.character(transect),
    Effort = 1,
    x = sampler_xy[, "X"],
    y = sampler_xy[, "Y"]
  ) |>
  select(Sample.Label, Effort, x, y)
 
# density surface model

dsm1 <- dsm(
  count ~ s(x, y),
  ddf.obj = m1,
  segment.data = segdata,
  observation.data = obsdata,
  family = quasipoisson(link = "log"),
  method = "REML",
  convert.units = 1)
 
summary(dsm1)

class(dsm1)
```
need abundance, density, detection ftn parameters

## Simulation 

region is already created

density should come from dsm1:

```{r}
# # grid for dsims density object
# 
# grid_for_density <- density_true@density.surface[[1]] |>
#   select(x, y) |>
#   mutate(area = 500 * 500)
# # DSM predicts abundance per grid cell
# 
# grid_for_density$Nhat <- predict(
#   dsm1,
#   newdata = grid_for_density,
#   off.set = grid_for_density$area
# )
#  
# # convert cell abundance to density
# 
# grid_for_density <- grid_for_density |>
#   mutate(Dhat = Nhat / area)
#  
# # fit a plain mgcv GAM that predicts density from x,y only
# 
# dsm_density_gam <- mgcv::gam(
#   Dhat ~ s(x, y),
#   data = grid_for_density,
#   family = gaussian(link = "log"),
#   method = "REML"
# )
#  
# # now this can be passed to dsims
# 
# dsm_density <- make.density(
#   region = region,
#   x.space = 500,
#   fitted.model = dsm_density_gam
# )
# 
# # plot(dsm_density, region, scale = 0.001)
```

```{r}
pred_grid <- density_true@density.surface[[1]] |>
  mutate(area = as.numeric(sf::st_area(geometry)))
 
pred_data <- sf::st_drop_geometry(pred_grid)
 
pred_grid$Nhat <- predict(
  dsm1,
  newdata = pred_data,
  off.set = pred_grid$area,
  type = "response"
)
 
dsm_surface <- pred_grid |>
  mutate(density = pmax(Nhat / area, .Machine$double.eps)) |>
  select(strata, density, x, y, geometry)
 
dsm_density <- make.density(
  region = region,
  x.space = 500,
  y.space = 500,
  density.surface = dsm_surface
)
 
plot(dsm_density, region, scale = 0.001)
```
 
If you only use the surface with `fixed.N = TRUE`, absolute scale does not matter; only relative cell weights matter. If you want a plot comparable to `density_true`, rescale for plotting/comparison, but do not confuse that with physical density:
 
```{r}
dsm_surface_plot <- dsm_surface |>
  mutate(density = density * sum(density_true@density.surface[[1]]$density) / sum(density))
```
 
One more review note: `coef(m1$ddf)` is on the link scale. Your half-normal scale estimate is approximately:
 
```{r}
sigma_hat <- exp(coef(m1$ddf)$scale["(Intercept)", "estimate"])
```
 
In my run that gives about `201.5`, matching the simulated `scale.param = 200`.



```{r}
# density <- density@density.surface[[1]]
# 
# fit.gam <- gam(density ~ s(x,y), data = density_true.data, family = gaussian(link="log"))

# Use the gam object to create a density object
gam.density <- make.density(region = region,
                            x.space = 500,
                            density.surface = dsm_surface)

plot(gam.density, region, scale = 0.001)
```

```{r}
# Define the population description
pop.desc <- make.population.description(region = region,
                                        density = gam.density,
                                        N = 1000,
                                        fixed.N = TRUE)
```

```{r}
# point.design <- make.design(region = region, 
#                             transect.type = "point",
#                             design = "systematic",
#                             spacing = 1500,
#                             edge.protocol = "minus",
#                             design.angle = 65,
#                             truncation = 750)
# 
# p.survey <- generate.transects(point.design)
# plot(region, p.survey)
```

```{r}
# Create the detectability description
detect <- make.detectability(key.function = "hn",
                             scale.param = 200,
                             truncation = 750)

# Plot the simulation detection functions
plot(detect, pop.desc)
```

```{r}
analyses <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             er.var = "P3",
                             criteria = "AIC")
```

```{r}
# sim.point <- make.simulation(reps = 1,
#                                 design = point.design,
#                                 population.description = pop.desc,
#                                 detectability = detect,
#                                 ds.analysis = analyses)
```

```{r}
# # Generate a single instance of a survey: a population, set of transects 
# # and the resulting distance data
# eg.point.survey <- run.survey(sim.point)
# 
# # Plot it to view a summary
# plot(eg.point.survey, region)
```

```{r}
# sim.point <- run.simulation(sim.point)
```

```{r}
# histogram.N.ests(sim.point)
```

```{r}
# ests_vals <- sim.point@results$individuals$N
#               
# ests <- as.data.frame.table(ests_vals) |>
#   pivot_wider(
#     id_cols = c(Var1, Var3), # Stratum and Run
#     names_from = Var2,       # Metric names (Estimate, se, cv, etc.)
#     values_from = Freq       # The data values
#   ) |>
#   slice_head(n = 99) |>
#   rename(region = Var1,
#          sim = Var3,
#          )
```

```{r}
# var(ests$Estimate)
```

#### Looped sim

```{r}
# Create grid of angles and pre-allocate the results vector
angles <- seq(0, 90, length.out = 100)
n_hat_i <- numeric(length(angles))

# Define static components outside the loop
detect <- make.detectability(
  key.function = "hn",
  scale.param = 200,
  truncation = 750
)

analyses <- make.ds.analysis(
  dfmodel = ~ 1,
  key = "hn",
  truncation = 750,
  er.var = "P3",
  criteria = "AIC"
)

# Loop over angles and store results
for (i in seq_along(angles)) {
  point.design <- make.design(
    region = region, 
    transect.type = "point",
    design = "systematic",
    spacing = 1500,
    edge.protocol = "minus",
    design.angle = angles[i],
    truncation = 750
  )
  
  sim.point <- make.simulation(
    reps = 1,
    design = point.design,
    population.description = pop.desc,
    detectability = detect,
    ds.analysis = analyses
  )
  
  sim.point <- run.simulation(sim.point)
  
  # Extract and store N for this iteration
  n_hat_i[i] <- sim.point@results[["individuals"]][["N"]][,,2][1]

}
```

```{r}
sd(n_hat_i)
```
















