---
title: "Variance Comparison"
format: pdf
---

Builds a test of our different variance options to see which compares best to true variance

- True variance comes from many iterations of:

  - Run a survey with a random start point on the 'true' population

  - Analyze with ds()

  - The variance of those many N_hat estimates from ds() is the true variance.

- Analytical variances, striplets, come from get_fit()

- Bootstrap estimates of N_hat from

1.  Create true density surface

-- outer loop start --

1.  Simulate true population from that surface

2.  Run a survey on that population (random start point)

3.  analyze with ds()

    1.  store N_hat

4.  Use ds() model results and survey data to create a dsm()

5.  simulate a population from that dsm()

6.  Calculate analytical variance estimators

-- inner loop start --

1.  Run bootstrap simulation models on estimated population to approx. striplets/analytical variances

-- inner loop end --

-- outer loop end --

# Packages

```{r}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
library(knitr)

library(future)
library(furrr)
```

# Import Functions

```{r}
source(here::here("Functions3.R"))
```

# Constants

```{r}
true_N <- 1000

points_or_lines = "point"

trunc_dist <- 300
scale_parameter <- 100
design_spacing <- 500

density_grid_spacing <- 500

design_angle <- 0
```

# Code

## Region

```{r}
region <- make_region(region_sf = NULL, 0, 5000, 0, 5000, "m")

plot(region)
```

## Density

```{r}
# my_hotspots <- list(
#   list(centre = c(2000, 8000), sigma = 800, amplitude = 1.2),
#   list(centre = c(5000, 2000), sigma = 6000, amplitude = 0.5),
#   list(centre = c(8000, 9000), sigma = 1000, amplitude = 2)
# )

my_hotspots <- list(
  list(centre = c(2500, 2500), sigma = 8000, amplitude = 0.1)
)

density_true <- make_hotspot_density(region, my_hotspots)
# plot(density_true, region, scale = 0.001)

density_sf <- density_true@density.surface[[1]]

ggplot(density_sf) +
  geom_sf(aes(fill = density), color = NA) +
  scale_fill_viridis_c(option = "viridis", name = "Density") +
  theme_bw() +
  labs(title = "Simulated Density Surface")
```

## Test Survey

```{r}
test_sim_pop <- generate_simulated_pop(
  region = region, 
  density_true = density_true, 
  scale_param = scale_parameter
)

test_population_description <- test_sim_pop$population_description
test_detectability <- test_sim_pop$detectability

test_survey_data_lines <- survey_data(
  region = region,
  realized_population = test_sim_pop$population,
  transect_type = points_or_lines
)

test_design <- test_survey_data_lines$design

test_analyses <- make.ds.analysis(dfmodel = ~ 1,
                             key = "hn",
                             truncation = 750,
                             criteria = "AIC")

test_sim <- make.simulation(reps = 3,
                                design = test_design,
                                population.description = test_population_description,
                                detectability = test_detectability,
                                ds.analysis = test_analyses)

# Generate a single instance of a survey: a population, set of transects
# and the resulting distance data
test_survey <- run.survey(test_sim)

# Plot it to view a summary
# plot(test_survey, region)
```

## Computing Speed

```{r}
# Detect cores and reserve 2 for system processes
available_cores <- availableCores()
workers_to_use <- max(1, available_cores - 2)

# Set the plan to use parallel workers
plan(multisession, workers = workers_to_use)
```

## Analysis Loop

```{r}
outer_loop_reps <- 3
bootstrap_reps <- 5

# Using future_map_dfr instead of a for-loop for speed
results <- future_map_dfr(1:outer_loop_reps, function(i) {
  
  ## 1. Generate population
  sim_pop <- generate_simulated_pop(
    region = region, 
    density_true = density_true
  )
  
  density <- sim_pop$density
  population <- sim_pop$population
  
  ## 2. Conduct survey
  survey_data_lines <- survey_data(
    region = region,
    realized_population = population,
    transect_type = points_or_lines
  )
  
  ## 3. Fit distance sampling model
  ds_results <- fit_ds(
    region = region,
    dist_data = survey_data_lines$dist_data,
    obsdata = survey_data_lines$obsdata,
    segdata = survey_data_lines$segdata,
    transect_type = points_or_lines,
    truncation = trunc_dist
  )
  
  ## 4. Fit density surface model
  dsm_results <- fit_dsm_and_surface(
    obsdata = survey_data_lines$obsdata,
    segdata = survey_data_lines$segdata,
    ds_model = ds_results$ds_model,
    region = region,
    transect_type = points_or_lines,
    x_space = density_grid_spacing,
    y_space = density_grid_spacing
  )
  
  # Failsafe in case nothing is observed and the DSM is NULL
  # FIXED: Added missing P2, P3, and boxlet columns to prevent summarise() crash
  if (is.null(dsm_results)) {
    return(tibble(
      rep = i,
      N_hat = ds_results$N_hat,
      se_ds_base = NA_real_,
      se_R2 = NA_real_, se_R3 = NA_real_,
      se_S1 = NA_real_, se_S2 = NA_real_,
      se_O1 = NA_real_, se_O2 = NA_real_,
      se_P2 = NA_real_, se_P3 = NA_real_, 
      se_striplet = NA_real_, se_boxlet = NA_real_,
      se_bootstrap = NA_real_,
      se_bootstrap_disc = NA_real_
    ))
  }
  
  ## 4.5 Build estimated population description for bootstraps
  # FIXED: Used dsm_surface which actually contains the calculated density column
  est_density <- make.density(
    region = region,
    x.space = density_grid_spacing,
    density.surface = dsm_results$dsm_surface 
  )
  
  pop_desc_est <- make.population.description(
    region = region,
    density = est_density,
    N = ds_results$N_hat,
    fixed.N = TRUE
  )
  
  ## 5. Calculate analytical variances
  spatial_variances <- striplet_boxlet(
    region = region,
    segdata = survey_data_lines$segdata,
    dsm_model = dsm_results$dsm_model,
    dsm_surface = dsm_results$dsm_surface,
    ds_results = ds_results,
    transect_type = points_or_lines,
    spacing = design_spacing,
    truncation = trunc_dist
  )
  
  # Extract analytical standard errors conditionally based on line or point 
  if (points_or_lines == "line") {
    R2_est <- ds_results$analytical_variances$se_N_R2
    R3_est <- ds_results$analytical_variances$se_N_R3
    S1_est <- ds_results$analytical_variances$se_N_S1
    S2_est <- ds_results$analytical_variances$se_N_S2
    O1_est <- ds_results$analytical_variances$se_N_O1
    O2_est <- ds_results$analytical_variances$se_N_O2
    P2_est <- NA_real_
    P3_est <- NA_real_
    striplet_est <- spatial_variances$se_N_striplet
    boxlet_est <- NA_real_
  } else {
    R2_est <- NA_real_
    R3_est <- NA_real_
    S1_est <- NA_real_
    S2_est <- NA_real_
    O1_est <- NA_real_
    O2_est <- NA_real_
    # FIXED: Was mapping to se_N_S1 erroneously
    P2_est <- ds_results$analytical_variances$se_N_P2 
    P3_est <- ds_results$analytical_variances$se_N_P3
    striplet_est <- NA_real_
    boxlet_est <- spatial_variances$se_N_boxlet
  }
  
  ## 6. Run standard bootstrap estimator
  bootstraps <- get_bootstrap(
    region = region,
    population_description = pop_desc_est,
    sigma_hat = ds_results$sigma_hat,
    transect_type = points_or_lines,
    reps = bootstrap_reps
  )
  
  bootstrap_est <- sd(bootstraps$N_hat, na.rm = TRUE)
  
  ## 7. Run discrete density bootstrap estimator
  bootstraps_disc <- get_bootstrap_disc_density(
    region = region,
    population_description = pop_desc_est,
    sigma_hat = ds_results$sigma_hat,
    transect_type = points_or_lines,
    reps = bootstrap_reps
  )
  
  bootstrap_disc_est <- sd(bootstraps_disc$N_hat, na.rm = TRUE)
  
  ## 8. Return a tibble/data.frame of this iteration's results
  # FIXED: Included all point/boxlet columns so the summarise chunk can find them
  tibble(
    rep                 = i,
    N_hat               = ds_results$N_hat,
    se_ds_base          = ds_results$se_ds, # Maps to the un-commented se_ds in Functions3
    se_R2               = R2_est,
    se_R3               = R3_est,
    se_S1               = S1_est,
    se_S2               = S2_est,
    se_O1               = O1_est,
    se_O2               = O2_est,
    se_P2               = P2_est,
    se_P3               = P3_est,
    se_striplet         = striplet_est,
    se_boxlet           = boxlet_est,
    se_bootstrap        = bootstrap_est,
    se_bootstrap_disc   = bootstrap_disc_est
  )
  
}, .options = furrr_options(seed = TRUE))
```

```{r}
results |>
  summarise(gold_standard = sd(N_hat, na.rm = TRUE),
            R2 = mean(se_R2, na.rm = TRUE),
            R3 = mean(se_R3, na.rm = TRUE),
            S1 = mean(se_S1, na.rm = TRUE),
            S2 = mean(se_S2, na.rm = TRUE),
            O1 = mean(se_O1, na.rm = TRUE),
            O2 = mean(se_O2, na.rm = TRUE),
            P2 = mean(se_P2, na.rm = TRUE),
            P3 = mean(se_P3, na.rm = TRUE),
            striplet = mean(se_striplet, na.rm = TRUE),
            boxlet = mean(se_boxlet, na.rm = TRUE),
            bootstrap = mean(se_bootstrap, na.rm = TRUE),
            bootstrap_disc = mean(se_bootstrap_disc, na.rm = TRUE)) |>
  kable(digits = 3)
```

Inflated striplet estimator variance, seems to be a product of the shifting density grid overlapping with values from shifting transects (500 and 750 will line up at 1500, 3000, etc) Improves when using a non-round value for design spacing.
