---
title: "Variance Comparison"
format: pdf
---

Builds a test of our different variance options to see which compares best to true variance

- True variance comes from many iterations of:

  - Run a survey with a random start point on the 'true' population

  - Analyze with ds()

  - The variance of those many N_hat estimates from ds() is the true variance.

- Analytical variances, striplets, come from get_fit()

- Bootstrap estimates of N_hat from

1.  Create true density surface

-- alternate outer loop start --

1.  Simulate true population from that surface

-- outer loop start --

1.  Run a survey on that population (random start point)

2.  analyze with ds()

    1.  store N_hat

3.  Use ds() model results and survey data to create a dsm()

4.  simulate a population from that dsm()

5.  Calculate analytical variance estimators

-- inner loop start --

1.  Run bootstrap simulation models on estimated population to approx. striplets/analytical variances

-- inner loop end --

-- outer loop end --

# Packages

```{r}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
library(knitr)
```

# Import Functions

```{r}
source(here::here("Functions.R"))
```

# Constants

```{r}
true_N <- 5000

points_or_lines = "line"

trunc_dist <- 750
scale_parameter <- 200
design_spacing <- 750

density_grid_spacing <- 500

design_angle <- 0
```

# Code

## Region

```{r}
region <- make_region(0, 10000, 0, 10000, "m")
plot(region)
```

## Density

```{r}
my_hotspots <- list(
  list(centre = c(2000, 8000), sigma = 800, amplitude = 1.2),
  list(centre = c(5000, 2000), sigma = 6000, amplitude = 0.5),
  list(centre = c(8000, 9000), sigma = 1000, amplitude = 2)
)

density_true <- make_hotspot_density(region, my_hotspots)
# plot(density_true, region, scale = 0.001)

density_sf <- density_true@density.surface[[1]]

ggplot(density_sf) +
  geom_sf(aes(fill = density), color = NA) +
  scale_fill_viridis_c(option = "viridis", name = "Density") +
  theme_bw() +
  labs(title = "Simulated Density Surface")
```

## True population

```{r}
sim_pop<- generate_simulated_pop(region)

# FIX: Map these to the exact names exported by the list in get_simulated_pop()
density <- sim_pop$density
population_description <- sim_pop$population_description
population <- sim_pop$population
detectability <- sim_pop$detectability
```

## Analysis Loop

```{r}
outer_loop_reps <- 10
bootstrap_reps <- 50

results <- data.frame(
  rep          = integer(outer_loop_reps),
  N_hat        = numeric(outer_loop_reps), # Usually good practice to keep the abundance estimate too
  se_R2        = numeric(outer_loop_reps),
  se_R3        = numeric(outer_loop_reps),
  se_S1        = numeric(outer_loop_reps),
  se_S2        = numeric(outer_loop_reps),
  se_O1        = numeric(outer_loop_reps),
  se_O2        = numeric(outer_loop_reps),
  se_striplet  = numeric(outer_loop_reps),
  se_bootstrap = numeric(outer_loop_reps)
)

for (i in 1:outer_loop_reps) {
  
  cat(sprintf("Running iteration %d of %d...\n", i, outer_loop_reps))
  
  ## Survey
  
  survey_data_lines <- generate_survey_data(
    region = region,
    realized_population = sim_pop$population,
    transect_type = points_or_lines
    )
  
  design <- survey_data_lines$design
  transects <- survey_data_lines$transects
  samplers <- transects@samplers
  
  ## Fit
  
  fit_lines <- get_fit(
    dist_data = survey_data_lines$dist_data,
    transects = survey_data_lines$transects,
    region = region,
    transect_type = points_or_lines
  )
  
  obs_data <- fit_lines$obsdata
  seg_data <- fit_lines$segdata
  dsm1 <- fit_lines$dsm
  density_surface <- fit_lines$density_surface
  
  R2_est <- fit_lines$analytical_variances$se_N_R2
  R3_est <- fit_lines$analytical_variances$se_N_R3
  S1_est <- fit_lines$analytical_variances$se_N_S1
  S2_est <- fit_lines$analytical_variances$se_N_S2
  O1_est <- fit_lines$analytical_variances$se_N_O1
  O2_est <- fit_lines$analytical_variances$se_N_O2
  striplet_est <- fit_lines$analytical_variances$se_N_striplet
  
  ## Bootstraps
  
  bootstraps <- get_bootstrap(
    region = region,
    population_description = fit_lines$population_description,
    sigma_hat = fit_lines$sigma_hat,
    transect_type = points_or_lines,
    reps = bootstrap_reps
  )
  
  bootstrap_est <- sd(bootstraps$N_hat)
  
  results$rep[i]          <- i
  results$N_hat[i]        <- fit_lines$N_hat # Storing the point estimate from the model for reference
  results$se_R2[i]        <- R2_est
  results$se_R3[i]        <- R3_est
  results$se_S1[i]        <- S1_est
  results$se_S2[i]        <- S2_est
  results$se_O1[i]        <- O1_est
  results$se_O2[i]        <- O2_est
  results$se_striplet[i]  <- striplet_est
  results$se_bootstrap[i] <- bootstrap_est
}
```

```{r}
results |>
  summarise(gold_standard = sd(N_hat),
            R2 = mean(se_R2),
            R3 = mean(se_R3),
            S1 = mean(se_S1),
            S2 = mean(se_S2),
            O1 = mean(se_O1),
            O2 = mean(se_O2),
            striplet = mean(se_striplet),
            bootstrap = mean(se_bootstrap)) |>
  kable(digits = 3)
```


