---
title: "Variance_Components"
format: html
---

1.  sets up basic region, density, population, and survey data

2.  runs get_fit

    1.  stores the dispersion parameter

3.  

# Packages

```{r}
library(tidyverse)
library(dsims)
library(Distance)
library(dssd)
library(dsm)
library(sf)
library(mgcv)
library(qrng) 
```

# Import Functions

```{r}
source(here::here("Functions.R"))
```

# Constants

```{r}
true_N <- 5000

points_or_lines = "line"

trunc_dist <- 750
scale_parameter <- 200
design_spacing <- 750
density_grid_spacing <- 500

design_angle <- 0
bootstrap_reps <- 100
```

# Code

## Region

```{r}
region <- make_region(region_sf = NULL, 0, 10000, 0, 10000, "m")
```

## Density

```{r}
my_hotspots <- list(
  list(centre = c(2000, 8000), sigma = 800, amplitude = 1.2),
  list(centre = c(5000, 2000), sigma = 6000, amplitude = 0.5),
  list(centre = c(8000, 9000), sigma = 1000, amplitude = 2)
)

density_true <- make_hotspot_density(region, my_hotspots)
```

## True population

```{r}
sim_pop<- generate_simulated_pop(region)

# FIX: Map these to the exact names exported by the list in get_simulated_pop()
density <- sim_pop$density
population_description <- sim_pop$population_description
population <- sim_pop$population
detectability <- sim_pop$detectability
```

## Survey

generates the iteration of survey data analogous to real data collection

```{r}
survey_data_lines <- generate_survey_data(
  region = region,
  realized_population = sim_pop$population,
  transect_type = points_or_lines
)

design <- survey_data_lines$design
transects <- survey_data_lines$transects
samplers <- transects@samplers
```

## Fit

```{r}
fit_lines <- get_fit(
  dist_data = survey_data_lines$dist_data,
  transects = survey_data_lines$transects,
  region = region,
  transect_type = points_or_lines
)

obs_data_base <- fit_lines$obsdata
seg_data <- fit_lines$segdata
dsm1 <- fit_lines$dsm
density_surface <- fit_lines$density_surface

# EXTRACT ESTIMATED PARAMETER: 
# Grab the estimated scale parameter from the m1 model 
# to hold detectability constant based on the base survey's fit.
m1_sigma_hat <- fit_lines$sigma_hat
```

## Bootstrap Settings

```{r}
# Set these to TRUE to let them vary, or FALSE to hold them constant across all runs
VARY_ANIMALS       <- F
VARY_TRANSECTS     <- T
VARY_DETECTABILITY <- T

# VARY_ANIMALS       <- FALSE
# VARY_TRANSECTS     <- FALSE
# VARY_DETECTABILITY <- FALSE

# Calculate probabilities and sizes for the animal placement generator
density_surface <- density_surface |> 
  mutate(N_expected = density * area)

prob_vec <- density_surface$N_expected / sum(density_surface$N_expected, na.rm = TRUE)

cell_dimensions <- sqrt(density_surface$area)

# Dummy population framework to appease dsims S4 requirements
dummy_density <- make.density(region = region, x.space = design_spacing, constant = 1)

dummy_pop_desc <- make.population.description(
  region = region, 
  density = dummy_density, 
  N = true_N, 
  fixed.N = TRUE
)

# detect <- make.detectability(
#   key.function = "hn", 
#   scale.param = scale_parameter, 
#   truncation = trunc_dist
# )

detect <- make.detectability(
  key.function = "hn", 
  scale.param = m1_sigma_hat, 
  truncation = trunc_dist
)

Region_Area <- as.numeric(sf::st_area(region@region))
```

Set up values for the case where we do not vary each element.

```{r}
# Freeze Animals?
if (!VARY_ANIMALS) {
  cell_counts <- as.vector(rmultinom(n = 1, size = true_N, prob = prob_vec))
  base_x <- rep(density_surface$x, cell_counts)
  base_y <- rep(density_surface$y, cell_counts)
  base_sizes <- rep(cell_dimensions, cell_counts)
  
  frozen_animal_x <- base_x + runif(true_N, min = -base_sizes/2, max = base_sizes/2)
  frozen_animal_y <- base_y + runif(true_N, min = -base_sizes/2, max = base_sizes/2)
}

# Freeze Transects?
if (!VARY_TRANSECTS) {
  frozen_transects <- generate.transects(design)
}

# Freeze Detectability?
if (!VARY_DETECTABILITY) {
  # Assign a permanent uniform deviate (hideability score) to every single animal (1 to N)
  set.seed(1209)
  frozen_det_rands <- runif(true_N, min = 0, max = 1)
}
```

```{r}
N_hat_results <- rep(NA, bootstrap_reps)

for (b in 1:bootstrap_reps) {
  
  # sim new animal positions
  if (VARY_ANIMALS) {
    cell_counts <- as.vector(rmultinom(n = 1, size = true_N, prob = prob_vec))
    base_x <- rep(density_surface$x, cell_counts)
    base_y <- rep(density_surface$y, cell_counts)
    base_sizes <- rep(cell_dimensions, cell_counts)
    
    animal_x <- base_x + runif(true_N, min = -base_sizes/2, max = base_sizes/2)
    animal_y <- base_y + runif(true_N, min = -base_sizes/2, max = base_sizes/2)
  } else { # use original animal positions
    animal_x <- frozen_animal_x
    animal_y <- frozen_animal_y
  }
  
  # rebuild population object with the correct animal positions
  realized_population <- dsims::generate.population(
    object = dummy_pop_desc, detectability = detect, region = region
  )
  native_pop <- realized_population@population
  native_pop$x <- animal_x
  native_pop$y <- animal_y
  realized_population@population <- as.data.frame(native_pop)
  
  
  # either re-place transects or use originals
  if (VARY_TRANSECTS) {
    current_transects <- generate.transects(design)
  } else {
    current_transects <- frozen_transects
  }
  
  
  # --- Step 3: Run the Survey ---
  if (VARY_DETECTABILITY) {
    # The 'Code as Written' option: dsims naturally varies binomial detection trials
    survey <- new("Survey.LT", 
                  population = realized_population, 
                  transect = current_transects, 
                  perp.truncation = trunc_dist)
    
    survey_run <- suppressWarnings(run.survey(survey, region = region))
    obs_data <- survey_run@dist.data
    
  } else {
    # The 'Frozen' option: Calculate geometry manually against the frozen uniform deviates
    pop_sf <- sf::st_as_sf(native_pop, coords = c("x", "y"))
    trans_sf <- current_transects@samplers
    
    # Fast distance matrix calculation
    dist_matrix <- sf::st_distance(pop_sf, trans_sf)
    units(dist_matrix) <- NULL # Drop sf units for fast matrix math
    
    # Find minimum distance to any transect for every animal
    closest_trans <- max.col(-dist_matrix, ties.method = "first")
    min_dists <- dist_matrix[cbind(1:nrow(dist_matrix), closest_trans)]
    
    # Filter only to animals within the truncation distance
    covered_idx <- which(min_dists <= trunc_dist)
    

    
    if (length(det_idx) > 0) {
      # Construct the data frame EXACTLY as ds() expects it
      obs_data <- data.frame(
        object = det_idx,                      # Unique ID for the animal
        Region.Label = region@region.name,     # Critical for extrapolation
        Area = Region_Area,                    # Critical for extrapolation
        Sample.Label = as.character(trans_sf$transect[closest_trans[det_idx]]),
        Effort = as.numeric(sf::st_length(trans_sf$geometry[closest_trans[det_idx]])),
        size = 1,
        distance = min_dists[det_idx]          # Must be named exactly 'distance'
      )
    } else {
      obs_data <- data.frame()
    }
  }
  
  
  # --- Step 4: Estimation ---
  # Because the stochasticity of detection is handled above, we just fit the model
  if (nrow(obs_data) > 0) {
    tryCatch({
      capture.output(
        m2 <- ds(data = obs_data, 
                 transect = points_or_lines, 
                 key = "hn", 
                 adjustment = NULL, 
                 truncation = trunc_dist, 
                 quiet = TRUE)
      )
      N_hat_results[b] <- as.numeric(m2$dht$individuals$N$Estimate)
    }, error = function(e) {
      N_hat_results[b] <- NA
    })
    
  } else {
    N_hat_results[b] <- 0
  }
}
```

```{r}
sim_results <- data.frame(
  replicate = seq_len(bootstrap_reps),
  N_hat = N_hat_results
)

sd(sim_results$N_hat, na.rm=TRUE)
VARY_ANIMALS
VARY_TRANSECTS
VARY_DETECTABILITY
```

With **all three** varied, 100 reps, sd(N_hat) was **87.83**

With only **detections** varied, 100 reps, sd(N_hat) was **28.00**

With only **animal position** varied, 100 reps, sd(N_hat) was **27.68**

With only **transect position** varied, 100 reps, sd(N_hat) was **25.03**

With **animals and transects** varied, 100 reps, sd(N_hat) was **34.72**

With **animals and detections** varied, 100 reps, sd(N_hat) was **67.01**

With **detections and transects** varied, 100 reps, sd(N_hat) was **61.85**
